---
title: "class09"
author: "Joshua Chevez"
date: "February 8, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Workshop: Cancer Cells

```{r}
fna.data <- "WisconsinCancer.csv"
read.csv(fna.data)
wisc.df <- read.csv(fna.data)
```

```{r}
head(wisc.df)
```

Note: last column x (all NA)
Check how many patients/measured variables
```{r}
#number of patients
nrow(wisc.df)
```

```{r}
#number of columns
ncol(wisc.df)
```

take columns 3 to 32
```{r}
wisc.data <- wisc.df[ , 3:32]
head(wisc.data)
```


set row names to patient ID numbers

```{r}
row.names(wisc.data) <- wisc.df$id
head(wisc.data)
```

How many malignant vs benign

```{r}
table(wisc.df$diagnosis)
```

How many features are "_mean" values
```{r}
colnames(wisc.data)
length(grep("_mean", colnames(wisc.data)))
```

```{r}
inds <- grep("_mean", colnames(wisc.data))
colnames(wisc.data)[inds]
```

#PC analysis


```{r}
colMeans(wisc.data)
```

Mean of every column (2) in the data
```{r}
round(apply(wisc.data,2,mean), 2)
```

Standard deviation of every column
```{r}
round(apply(wisc.data,2,sd), 2)
```
#PCA
Different so needs to be scaled

```{r}
wisc.pr <- prcomp(wisc.data, scale. = TRUE)
summary(wisc.pr)
biplot(wisc.pr)
#biplot doesn't help


```

Make a useful plot; Each dot represents a patient.
```{r}
plot(wisc.pr$x[, 1], wisc.pr$x[, 2])
```


Lets color in the diagnosis; malignant in red (black is the first color for the first level, red is the default color for the second vector)
```{r}
wisc.df$diagnosis
plot(wisc.pr$x[, 1], wisc.pr$x[, 2], col = wisc.df$diagnosis)
```

Scree plot; Variance = sdev^2; total variance = sum(variance)
```{r}
variance <- wisc.pr$sdev^2
variance
pve <- round((variance/sum(variance)) * 100, 2)
```

```{r}
plot(pve, type = "o")
```

Barplot; remove and add an axis and add values
```{r}
barplot(pve, axes = FALSE, names.arg = paste("PC", 1:length(pve)), las = 2, ylab = "Proportion of Variance (%)")
axis(2, round(pve))
```

#Clustering using ward.d2 heirarchial clustering for our PCA

For hclust we need a distance matrix from our PCA results (wisc.pr$x)
```{r}
d <- dist(wisc.pr$x[, 1:2])
hc <- hclust(d, method = "ward.D2")
plot(hc)
```

Cut our tree into 3 groups

```{r}
grp3 <- cutree(hc, k = 3)
table(grp3)
```

Our PCA plot colored by cluster
```{r}
plot(wisc.pr$x[, 1], wisc.pr$x[, 2], xlab = "PC1", ylab = "PC2", col = grp3)
```

Cross tabulation
```{r}
diagnosis <- wisc.df$diagnosis == "M"
table(grp3, diagnosis)
```

Same thing:
```{r}
table(grp3, wisc.df$diagnosis)
```

```{r}
new <- read.csv("new_samples.csv")
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[, 1:2], col=wisc.df$diagnosis)
points(npc[, 1], npc[, 2], col = c("green", "blue"), pch = 15, cex = 3)
```














