---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(bio3d)
```


Download structure from pdb database "1hsg"
get.pbd gets the file and puts it into the working directory

Name pdb code so that it is easily changable
```{r}
pdb.code <- "1hsg"
file.name <- get.pdb(pdb.code)

```

Extract the protein only segment of the pdb entry and write out new pdb format file. Do the same for the bound ligand


```{r}
hiv <- read.pdb(file.name)
hiv
```

Protein extraction

```{r}
prot <- trim.pdb(hiv, "protein")
prot
```
Make any name up for file (not hard coded and can be easily changed)

```{r}
prot.filename <- paste(pdb.code, "_protein.pdb", sep = "")
write.pdb(prot, file = prot.filename)
```



Same thing for ligand

```{r}
lig <- trim.pdb(hiv, "ligand")
write.pdb(lig, file = paste(pdb.code, "_ligand.pdb", sep = ""))
```

```{r}
lig
```

##Convert docking results for viewing in VMD
14 models (answers)

```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```


Calculating distance/similarity
```{r}
# res <- read.pdb("all.pdbqt", multi=TRUE)
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```

#Hen egg white lysozyme
```{r}
library(bio3d)
pdb <- read.pdb("1HEL")
modes <- nma(pdb)
plot(modes, sse=pdb)
```

```{r}
mktrj(modes, mode=7, file="nma_7.pdb")

```














